Accelerated evolution of convective simulations

### 2018-03-01 tex/bvp_paper.tex
updates final paragraph and all figure captions; finalizing edits tomorrow

\begin{figure}[t]
\includegraphics[width=\textwidth]{./figs/time_trace.png}
\caption{A time trace is shown of kinetic energy (black) and mean temperature (blue)  vs. time 
for an SE run at $S = 10^5$ (a). The mean evolved values of kinetic energy and mean temperature
are denoted by the horizontal dashed black and blue lines. 
By taking an average of the fluxes early in the
simulation (b), and appropriately using the AE method described here, the evolution
of these atmospheres can be greatly accelerated. The time trace of such an accelerated
atmosphere, which uses these methods, is shown in (c).  The y-axes and x-axes are
scaled identically as in (a).  Here, the AE method is used three times, marked by
the numbered arrows. The fluxes averaged over the green region which begins at
the vertical dashed line shortly after AE 3 are shown in (c) are shown
in (d), and they are nearly in the fully evolved state, sampled over a similarly
long region that occurs much later in the simulation in (a).  The difference between
the fluxes in the AE and SE ssolutions is shown in (e), and it is small. The dashed
vertical line in (a) denotes the simulation time at which evolved flux measurements
begin to be taken in (c).
\label{fig:time_trace} }
\end{figure}

@@

The AE method converges atmospheeres from extreme flux disequilibrium states 
(Fig. \ref{fig:time_trace}b) on short dynamical timescales (Fig. \ref{fig:time_trace}c)
to a converged state (Fig. \ref{fig:time_trace}d) whose flux balance is very
close to the balance present in the corresponding SE solution (Fig. \ref{fig:time_trace}e).
While the application of the AE method advances solutions toward the proper solution,
we find that multiple applications of the method is the best means of achieving converged
solutions quickly.

@@

\begin{figure}[t]
\includegraphics[width=\textwidth]{./figs/parameter_space_comparison.png}
\caption{Volume- and time-averaged measurements of the Nusselt number, the
RMS Reynolds number, and the mean temperature for AE runs are shown in (a)-(c).
Symbols are located at the mean value of
each measurement, vertical lines represent the standard deviation of the measurement,
which shows natural variation over the
averaging window, and error bars represent the change in the mean value over the averaging window,
such that very large error bars represent solutions which are not fully converged when averaging
begins.
(a) The Nusselt number scales as Ra$^{1/5}$, and above S$\geq 10^{3+2/3}$,
simulations exhibit oscillating plume structures whose heat transport varies over time.  
(b) The RMS Reynolds number, which measures the level of turbulence in the evolved solution, scales as
Ra$^{0.45}$. (c) Temperature, with its value at the upper boundary removed, is shown,
and scales as Ra$^{-1/5}$.
Relative error for (d) Nu, (e) Re, and (f) $T$ are shown between measurements 
taken in AE solutions and SE solutions.
The greyed area of the plot indicates the region in which only AE runs were
carried out due to computational expense. The run at $S = 10^5$ marked as a
black square is examined in more details in Figs. \ref{fig:time_trace},
\ref{fig:temp_comparison}, \& \ref{fig:pdf_comparison}.
\label{fig:parameter_space_comparison} }
\end{figure}

@@

@@

\begin{figure}[t]
\includegraphics[width=\textwidth]{./figs/temp_comparison.png}
\caption{Comparisons of the evolved thermodynamic states of an AE and SE run
at $S = 10^{5}$ are shown.  (a) Evolved horizontally- and time-averaged 
temperature profiles, as a function of height.
(b) Probability Distribution Functions (PDFs) and their integrated
Cumulative Distribution Functions (CDFs)
of point-by-point measurements of the temperature field.
(c) The percentage difference between the mean temperature profiles as a function of height.
(d) The value of the Kolmogorov-Smirnov (KS) statistic, or the difference between
the AE and SE CDFs, as a function of temperature.
$T$ is sampled every 0.1 time units for 500 total time units,
and was interpolated onto an evenly spaced grid before sampling.
The difference between the mean profiles is very small O(0.5\%), but this small
difference results in a large difference between the two temperature CDFs near the values
of the temperature mode.  The spread of temperature around the mode, or the fluctuations that
drive convection, are nearly identical between the two runs.
\label{fig:temp_comparison} }
\end{figure}

@@

\begin{figure}[t]
\includegraphics[width=\textwidth]{./figs/pdf_comparison.png}
\caption{Probability distribution functions of (a) the vertical velocity, (b) the horizontal velocity, and (c) nonlinear
convective transport are shown for a 2D runs achieved through the SE and AE methods
at $S = 10^{5}$.  Flows are sampled every 0.1 time units for 500 total time units,
and flows are interpolated onto an evenly spaced grid before sampling.
The cumulative distribution function is overplotted on each plot. 
\label{fig:pdf_comparison} }
\end{figure}

@@

A long-term goal would be to extend AE in a robust way such that it can be
incorporated into stellar structure codes. Current state of the art stellar structure
models are 1D solves that parameterize convection through widely accepted models
like Mixing Length Theory \cite{paxton&all2011}. Recently, researches have begun
efforts to couple evolution of stellar structure on significant thermal timescales
while resolving convective dynamics.  A first attempt to do this was through the
use of fully nonlinear implicit timestepping methods 
\cite{viallet&all2011, viallet&all2013, viallet&all2016},
but the emergence of unrealistic dynamics when timesteps are so large that they
don't properly resolve advective motions has made this an unlikely means of
evolving systems for long times. Other recent work, 
such as 3-2-1D \cite{arnett&all2015} convection, which uses information in
evolved 3D dynamical solutions to evolve the 1D underlying thermodynamic profile,
and which is similar to the method we presented here, shows promise.  Some other
work pairing 1D solves with fully nonlinear 3D solutions has been conducted
\cite{cristini&all2016}, but often in more complicated systems than those we examine here. 
Through small improvements and extensions, the AE method here can be improved
and verified until it is capable of handling the full physics present in nonlinear
stellar evolution.

>>>>

\begin{figure*}[t!]
\includegraphics[width=\textwidth]{./figs/time_trace.png}
\caption{(a) Kinetic energy (black) and mean temperature (blue)  vs. time are shown
for a SE run at $S = 10^5$. The mean evolved values of kinetic energy and mean temperature,
averaged over the time shaded in green,
are denoted by the horizontal dashed lines. (b) The time- and horizontally-averaged
flux profiles are shown for the times highlighted in orange in (a).
(c) The same quantities as in (a) are shown, but for AE at the same parameters.
The axes are scaled identically in (a) and (c), and the AE method is used three times, marked by
the numbered arrows in (c). The fluxes averaged over the green shaded region of (c)
are shown in (d). The difference between
the fluxes in the AE and SE solutions is shown in (e). The dashed
vertical line in (a) and (c) denotes the simulation time at which evolved measurements
are started for the AE case.
\label{fig:time_trace} }
\end{figure*}

@@

The AE method converges atmospheres from extreme flux disequilibrium states 
(Fig. \ref{fig:time_trace}b) on short dynamical timescales (Fig. \ref{fig:time_trace}c)
to a converged state (Fig. \ref{fig:time_trace}d) whose flux balance is very
close to the balance present in the corresponding SE solution (Fig. \ref{fig:time_trace}e).
While the application of the AE method advances solutions toward the proper solution,
we find that multiple applications of the method is the best means of achieving converged
solutions quickly.

@@

\begin{figure}[b]
\includegraphics[width=\textwidth]{./figs/parameter_space_comparison.png}
\caption{Volume- and time-averaged measurements of the Nusselt number (Nu), the
RMS Reynolds number (Re), and the mean temperature ($\angles{T}$) for AE runs are shown in (a)-(c).
Symbols are located at the mean value of
each measurement.  Vertical lines represent the standard deviation of the measurement,
and quantify natural variation over the
averaging window. Error bars represent the change in the mean value over the averaging window,
such that very large error bars represent solutions which are not fully converged when averaging
begins.
(a) Nu scales as Ra$^{1/5}$; above S$\geq 10^{3+2/3}$,
simulations exhibit oscillating plume structures whose heat transport fluctuates over time.  
(b) Re, which measures the level of turbulence in the evolved solution, scales as
Ra$^{0.45}$. (c) The difference between $\angles{T}$ and the value of $T$ at $z = +1/2$
is shown, and this quantity scales as Ra$^{-1/5}$, the inverse of Nu.
Relative error for measurements of (d) Nu, (e) Re, and (f) $T$ between 
AE solutions and SE solutions are shown.
The greyed area of the plots indicates the region in which only AE runs were
carried out due to computational expense. The run at $S = 10^5$ marked as a
black square is examined in more details in Figs. \ref{fig:time_trace},
\ref{fig:temp_comparison}, \& \ref{fig:pdf_comparison}.
\label{fig:parameter_space_comparison} }
\end{figure}

@@

\begin{figure}[t]
\includegraphics[width=\textwidth]{./figs/temp_comparison.png}
\caption{Comparisons of the evolved thermodynamic states of an AE and SE run
at $S = 10^{5}$ are shown.  (a) Evolved horizontally- and time-averaged 
temperature profiles, as a function of height.
(b) Probability Distribution Functions (PDFs) and their integrated
Cumulative Distribution Functions (CDFs)
of point-by-point measurements of the temperature field.
$T$ is sampled every 0.1 time units for 500 total time units,
and was interpolated onto an evenly spaced grid before sampling.
(c) The percentage difference between the mean temperature profiles as a function of height.
The difference between the mean profiles is very small, O(0.5\%).
(d) The value of the Kolmogorov-Smirnov (KS) statistic, or the difference between
the AE and SE CDFs, as a function of temperature.
The small
difference in the interior temperature 
results in a large difference between the two temperature CDFs near the values
of the temperature modes.  The spread of temperature around the modes, which includes the 
fluctuations that drive convection, are nearly identical between the two runs.
\label{fig:temp_comparison} }
\end{figure}

@@

@@

\begin{figure}[t]
\includegraphics[width=\textwidth]{./figs/pdf_comparison.png}
\caption{Probability distribution functions (PDFs) of (a) the vertical velocity, (b) the horizontal velocity, and (c) nonlinear
convective transport are shown for 2D runs achieved through SE (blue) and AE (red)
at $S = 10^{5}$.  Flows are sampled every 0.1 time units for 500 total time units,
and are first interpolated onto an evenly spaced grid.
The cumulative distribution function (CDF) is overplotted for each PDF. 
(d-f) The KS statistics, or the values of CDF$_{\text{AE}}$ - CDF$_{\text{SE}}$,
are shown for the related distributions, and solid lines indicated positive values
while dashed lines are negative values. Unlike the temperature distributions in
Fig. \ref{fig:temp_comparison}, these distributions, particularly the vertical
velocity and transport, show very good agreement and small values of the KS statistic.
\label{fig:pdf_comparison} }
\end{figure}

@@

An eventual use case of AE is in robustly replacing Mixing Length Theory (MLT)
in stellar evolution codes. Current state of the art stellar structure
models are the solutions of 1D simulations which parameterize convection 
using MLT \cite{paxton&all2011}. Recent work has taken the first steps towards
understanding how to simultaneously resolve convective motions while evolving
systems through many thermal timescales, as is required in the evolution of stars.
We now know that implicit timestepping methods, while not technically bound by
CFL constraints, must resolve the convective motions for stability
\cite{viallet&all2011, viallet&all2013, viallet&all2016}, and are thus
not an ideal solution to achiving long-term system evolution.
Recently, efforts have begun to improve 1D models of stellar
convection by projecting the results of 3D simulations into 1D
to more properly parameterize convection \cite{arnett&all2015, cristini&all2016}.
This is precisely an extension of the AE that we present here,  and through the
careful addition and testing of new physics, AE shows great potential for 
improving our understanding of convection in general, and stellar evolution models
in particular.
